\section{Introduction}
%uvodne kecy
%   
World is getting smaller every day, as new technologies constantly make communication and traveling faster and more effective then yesterday. Road network, Internet and many other networks are becoming more evolved and denser which also brings along new problems. In order to take advantage of such huge networks, we must have efficient algorithms that operate on these networks and give us answers to many questions. Among many others, one that we take great interest in is the question: ``What is the shortest path from place $x$ to place $y$''? \\

In different networks, this question can make different sense. In the road network, we would like to obtain a sequence of intersections we have to visit in order to reach our destination, driving the shortest possible time (or the smallest possible distance) . GPS devices and the likes of Google maps have to deal with this problem. In the case of Internet network, we might be interested in the shortest path to a destination computer in terms of router hops. In a network of social acquaintances, the smallest number of persons connecting us with e.g. Mark Knopfler could be expressed as a shortest path problem. Many problems in artificial intelligence (e.g. planning of actions to obatain some goal) can be expressed (or include) as a shortest path problem. \\

The importance of finding a shortest path in a graph is also obvious from the amount of algorithms and approaches we have nowadays to tackle this problem. Many traditional methods (like Dijkstra's algorithm) have many variations %TODO - check 
for different settings. 

%applications

%tradicne pristupy

%organization of the thesis
